---
title: "Data Manipulation"
output: github_document
---

```{r tidyverse, message = FALSE}
library(tidyverse)
library(ggridges)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit plot object

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))
  
weather_plot + geom_point()
```

## Advanced Scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


Placement of aesthetic matters when using multiple geom functions

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

Facet - Create multi layer plots
Aesthetic elements per geom element

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5, size = .3) +
  geom_smooth(se = FALSE, size = 1.5) + 
  facet_grid(. ~ name)

#facet_grid(row ~ column)
```


Combine elements - create new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


## Small Notes

Geom Types

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

Neat Geom! 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = 0.3)
```


## Univariate Plots

#### Histograms

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Addition of color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

#### Density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .4, adjust = .5)

# adjust smooths more or less
```

#### Box-plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

### Trendy Plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```

Ridge plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and Embed

Save a scatterplot

```{r}
weather_plot = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("weather_plot.pdf", weather_plot)

# ggsave("./results/weather_plot.pdf", weather_plot, width  = 8, height = 5)
```

Embedding

```{r}
weather_plot
```

```{r, fig.width = 12, fig.height = 6}
weather_plot
```



