---
title: "Data Manipulation"
subtitle: Part II
output: github_document
---

```{r tidyverse, message = FALSE}
library(tidyverse)
library(patchwork)
```

## Load Weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Starting Plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)
```


## Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
```

## Scales

#### X and Y Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    #trans = "log"
    position = "right"
  )
```


#### Color Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 200))
```

Viridis Package

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE)
```


## Themes

#### Shift Legend

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

#### Theme_bw()

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) +
  theme_bw()
```

#### Theme_minimal()

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE) +
  theme_minimal() +
#ggthemes::****
  theme(legend.position = "bottom")
```

If using a theme, place theme specifications after selected theme


## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "virirdis"
)

scale_colour_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```


## Data arguments in `geom`

```{r}
central_park =
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki =
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

## Patchwork

remember faceting

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

Multipanel plots without faceting

```{r}
tmax_tmin_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) +
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

tmax_tmin_plot + prcp_dens_p + tmax_date_p

tmax_tmin_plot + (prcp_dens_p + tmax_date_p)

tmax_tmin_plot / (prcp_dens_p + tmax_date_p)

(prcp_dens_p + tmax_date_p) / tmax_tmin_plot
```

## Data Manipulation

Control Factors. 

```{r}
weather_df %>% 
  mutate(
    name = factor(name), 
    name = forcats::fct_relevel(name, c("Waikiki_HA"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```


Densities for tmin and tmac simultaneously

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5)
```

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

## Revisit pups

Data from FAS study

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate( sex = recode(sex, `1` = "male", `2` = "female"))

litter_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number")



fas_data %>% 
  select(dose, day_of_tx, starts_with("pd_")) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day"
  ) %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_relevel(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
```

















